---
title: "190919_inclass"
author: "Gavin Ko"
date: "9/19/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library and load datasets}

library(tidyverse)
litters_data = read.csv("FAS_litters.csv")

pups_data = read.csv("FAS_pups.csv")

```
select and mutate is useful in selecting those interested variable.

```{r janitor for label names cleaning}
litters_data = janitor::clean_names(litters_data)
pull(litters_data, group)
```

R doesn't know what's inside a dataset. We need to tell them where to look.

Now then, practice select.

First one variable is the dataset, from the second you can start sorting whatever you want.

```{r selection}
select(litters_data, litter_number, group, gd0_weight)

```

everything() serve as the function to list everything else then designated.
```{r selection for all}
select(litters_data, litter_number, everything() )
```

if you put minus sign in front of a variable, it means that you're excluding it from the dataset. 
```{r selection except}
select(litters_data, -litter_number)
```

You can even indicate the coverage like excel format.
```{r selection for some}
select(litters_data, litter_number, gd0_weight:pups_born_alive)
```

do notice that the selection function dosen't change the original dataset, unless you assign a new one.

You can also rename the labels during the process. Put the new name you want on the left.
If you only put rename inside, it doesn't change the data sequence.
```{r}
select(litters_data, GROuP = group, litter_number)
select(litters_data, GROUP = group)
```


Secondly, let's move on to filter. Filter serves as conditional selection.

```{r}
filter(litters_data, group == "Con7")
```

With filter, you can give all kinds of logical function.
Simply add the multiple filterations by adding it after comma.
Second line of writing or condition is better, since "|" cause some compiling trouble sometimes.

```{r mischellaneous}
filter(litters_data, pups_born_alive < 6 , group == "Con7")
filter(litters_data, group %in% c("Con7", "Con8") )
filter(litters_data, pups_born_alive <= 6, pups_born_alive >= 4)
filter(litters_data, gd0_weight + gd18_weight >= 70)
#don't do this :filter(litters_data, !is.na(gd0_weight))
droup_na(litters_data) # droup all rows with any single na
drop_na(litters_data, gd0_weight) #only focus on the selected column
```

Alright, let's do mutate right now.

Note that mutate is inreversable changes.
```{r mutate}
mutate(litters_data, 
       wt_gain = gd18_weight - gd0_weight, #add new column
       group = str_to_lower(group) #change the variable value; character to lower case
       )
```

Finally, go on to arrange. It's most of the time for double checking the data if there's anything unexpected or wrong. desc() indicates reverse order.
```{r arrange}
arrange(litters_data, desc( pups_born_alive) )

arrange(litters_data, pups_born_alive, gd0weight)
#multiple clusters, what goes first, filtered the first.

```


